---
title: "Analytics"
output: md_document
params:
  healthRisk: "AWD"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
knitr::opts_chunk$set(echo = F)

map <- readRDS("data/gadm36_NOR_1_sp.rds")
  #map <- rmapshaper::ms_simplify(map,keep = 0.5)
map <- broom::tidy(x=map,region="NAME_1")
setDT(map)
unique(map$id)
#saveRDS(map,"data/map_supervisor.RDS")


WeekC <- function (date = lubridate::today()) 
{
    wk <- as.numeric(format.Date(date, "%V"))
    wk <- formatC(wk, flag = "0", width = 2)
    return(wk)
}

Year <- function (date = lubridate::today()) 
{
    yr <- as.numeric(format.Date(date, "%G"))
    return(yr)
}

MonthC <- function (date = lubridate::today()) 
{
    m <- as.numeric(format.Date(date, "%m"))
    m <- formatC(m, flag = "0", width = 2)
    return(m)
}

Month <- function (date = lubridate::today()) 
{
    yr <- as.numeric(format.Date(date, "%m"))
    return(yr)
}

YearMonth <- function(date){
  return(sprintf("%s-%s", Year(date), MonthC(date)))
}

YearWeek <- function(date){
  return(sprintf("%s-%s", Year(date), WeekC(date)))
}

YearWeekN <- function(date){
  return(as.numeric(sprintf("%s%s", Year(date), WeekC(date))))
}

dr <- readxl::read_excel("data/casereports-2018-September-28 (1).xlsx")
dc <- readxl::read_excel("data/datacollectors-2018-September-28 (1).xlsx")

setDT(dr)
setDT(dc)


setnames(dr,"Data Collector","datacollector")
setnames(dr,"Males < 5","mu5")
setnames(dr,"Males ≥ 5","mo5")
setnames(dr,"Females < 5","fu5")
setnames(dr,"Females ≥ 5","fo5")

dr <- rbind(dr,dr,dr,dr,dr)

dr[,n:=floor(1:.N/2),by=datacollector]
dr[,Date:=lubridate::today()-n]
dr[,n:=NULL]

dr <- dr[`Health Risk`==params$healthRisk]
dr[,`Health Risk`:=NULL]
setnames(dr,"Lat. / Long.","latlong")

dr[,Region:=NULL]
dr[,District:=NULL]
dr[,Village:=NULL]
dr[,latlong:=NULL]

locs <- dc[,c("Display Name","Lat. / Long.","Region","District","Village")]
setnames(locs,c("datacollector","latlong","region","district","village"))


minDate <- min(dr$Date)
maxDate <- max(dr$Date)
weeks <- data.table(Date=seq.Date(minDate,maxDate,1))
weeks[,rollingWeek:=floor(as.numeric(difftime(max(Date),Date,units="days"))/7)]
weeks[,epiWeek:=YearWeek(Date)]
weeks[,epiWeekN := YearWeekN(Date)]
weeks[,rollingMinDate:=min(Date),by=rollingWeek]
weeks[,rollingMaxDate:=max(Date),by=rollingWeek]
weeks[,epiMinDate:=min(Date),by=epiWeek]
weeks[,epiMaxDate:=max(Date),by=epiWeek]
weeks[,rollingx:=max(rollingWeek)-rollingWeek]
weeks[,epix:=max(epiWeekN)-epiWeekN]
weeks[,epiMonth:=YearMonth(Date)]

weeksNoDate <- copy(weeks)
weeksNoDate[,Date:=NULL]
weeksNoDate <- unique(weeksNoDate)

skeleton <- vector("list",length=nrow(locs))
for(i in 1:nrow(locs)){
  skeleton[[i]] <- copy(weeks)
  skeleton[[i]][,datacollector:=locs$datacollector[i]]
}
skeleton <- rbindlist(skeleton)
nrow(skeleton)
skeleton <- merge(skeleton,locs,by=c("datacollector"),all.x=T,allow.cartesian = T)
nrow(skeleton)
d <- merge(skeleton,dr,by=c("Date","datacollector"),all.x=T,allow.cartesian = T)
nrow(d)
d[,isMessage:=ifelse(is.na(Time),0,1)]
xtabs(~d$isMessage)
d[,isMessageValid:=isMessage]
d[isMessage==1,isMessageValid:=sample(c(1,0),size=.N,replace=T,prob=c(0.7,0.3))]

d[is.na(mu5),mu5:=0]
d[is.na(mo5),mo5:=0]
d[is.na(fu5),fu5:=0]
d[is.na(fo5),fo5:=0]

dWide <- d
n <- names(d)[!names(d) %in% c("mu5","mo5","fu5","fo5")]
d <- melt.data.table(d,id.vars=n)

d[,sex:="Male"]
d[variable %in% c("fu5","fo5"),sex:="Female"]

d[,age:="Age <5"]
d[variable %in% c("mo5","fo5"),age:="Age 5+"]

```

## Epicurve by week

Here we display a weekly `epicurve` (the epidemiological term for a time series graph showing the number of reported cases on the y-axis and time on the x-axis).

Important to note:

- We display `year-isoweek` on the x-axis
- Weeks with zero cases must be displayed

```{r message=FALSE, warning=FALSE}
toplotNational <- d[,.(n=sum(value)),keyby=.(epiWeek,age,sex)]
toplotNational[,level:="all"]
toplotNational[,geo:="all"]
toplotRegion <- d[,.(n=sum(value)),keyby=.(region,epiWeek,age,sex)]
toplotRegion[,level:="region"]
setnames(toplotRegion,"region","geo")
toplotDistrict <- d[,.(n=sum(value)),keyby=.(district,epiWeek,age,sex)]
toplotDistrict[,level:="district"]
setnames(toplotDistrict,"district","geo")
toplotVillage <- d[,.(n=sum(value)),keyby=.(village,epiWeek,age,sex)]
toplotVillage[,level:="village"]
setnames(toplotVillage,"village","geo")
tosave1 <- rbind(toplotNational,toplotRegion,toplotDistrict,toplotVillage)
tosave2 <- tosave1[,.(n=sum(n)),keyby=.(level,geo,epiWeek,age)]
tosave2[,sex:="All"]
tosave3 <- tosave1[,.(n=sum(n)),keyby=.(level,geo,epiWeek,sex)]
tosave3[,age:="All"]
tosave4 <- tosave1[,.(n=sum(n)),keyby=.(level,geo,epiWeek)]
tosave4[,sex:="All"]
tosave4[,age:="All"]
tosave <- rbind(tosave4,tosave3,tosave2,tosave1)
tosave[,pop:=10000]

readr::write_lines(jsonlite::toJSON(tosave),path="json/epicurve-by-week.json")

q <- ggplot(tosave4,aes(x=epiWeek,y=n))
q <- q + geom_col(alpha=0.7)
q <- q + labs(title=sprintf("Weekly epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
q <- q + scale_x_discrete("Epiweek")
q <- q + scale_y_continuous("Number of reported cases")
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```

## Epicurve by day

Here we display a daily `epicurve`.

Important to note:

- Unclear the best way to display date on the x-axis
- Days with zero cases must be displayed

```{r message=FALSE, warning=FALSE}
toplotNational <- d[,.(n=sum(value)),keyby=.(Date,age,sex)]
toplotNational[,level:="all"]
toplotNational[,geo:="all"]
toplotRegion <- d[,.(n=sum(value)),keyby=.(region,Date,age,sex)]
toplotRegion[,level:="region"]
setnames(toplotRegion,"region","geo")
toplotDistrict <- d[,.(n=sum(value)),keyby=.(district,Date,age,sex)]
toplotDistrict[,level:="district"]
setnames(toplotDistrict,"district","geo")
toplotVillage <- d[,.(n=sum(value)),keyby=.(village,Date,age,sex)]
toplotVillage[,level:="village"]
setnames(toplotVillage,"village","geo")
tosave1 <- rbind(toplotNational,toplotRegion,toplotDistrict,toplotVillage)
tosave2 <- tosave1[,.(n=sum(n)),keyby=.(level,geo,Date,age)]
tosave2[,sex:="All"]
tosave3 <- tosave1[,.(n=sum(n)),keyby=.(level,geo,Date,sex)]
tosave3[,age:="All"]
tosave4 <- tosave1[,.(n=sum(n)),keyby=.(level,geo,Date)]
tosave4[,sex:="All"]
tosave4[,age:="All"]
tosave <- rbind(tosave4,tosave3,tosave2,tosave1)
tosave[,pop:=10000]

readr::write_lines(jsonlite::toJSON(tosave),path="json/epicurve-by-day.json")

q <- ggplot(tosave4,aes(x=Date,y=n))
q <- q + geom_col(alpha=0.7,width=0.5)
q <- q + labs(title=sprintf("Daily epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
#q <- q + scale_x_discrete("Epiweek")
q <- q + scale_y_continuous("Number of reported cases")
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```

## Epicurve by week dodged by age

Here we display a weekly `epicurve` with two columns for each week, showing the ages side-by-side.

Important to note:

- We display `year-isoweek` on the x-axis
- Weeks with zero cases must be displayed

```{r message=FALSE, warning=FALSE}
toplotNational <- d[,.(n=sum(value)),keyby=.(age,epiWeek)]
toplotNational[,level:="all"]
toplotNational[,geo:="all"]
toplotDistrict <- d[,.(n=sum(value)),keyby=.(age,district,epiWeek)]
toplotDistrict[,level:="district"]
setnames(toplotDistrict,"district","geo")
toplotVillage <- d[,.(n=sum(value)),keyby=.(age,village,epiWeek)]
toplotVillage[,level:="village"]
setnames(toplotVillage,"village","geo")
tosave <- rbind(toplotNational,toplotDistrict,toplotVillage)
tosave[,pop:=10000]

#readr::write_lines(jsonlite::toJSON(tosave),path="json/epicurve-by-week-dodged-by-age.json")

q <- ggplot(toplotNational,aes(x=epiWeek,y=n,fill=age))
q <- q + geom_col(pos="dodge")
q <- q + labs(title=sprintf("Weekly epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
q <- q + scale_x_discrete("Epiweek")
q <- q + scale_y_continuous("Number of reported cases")
q <- q + scale_fill_brewer("Age",palette="Set1")
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```

## Epicurve by day dodged by age

Here we display a daily `epicurve` with two columns for each day, showing the ages side-by-side.

Important to note:

- Unclear the best way to display date on the x-axis
- Days with zero cases must be displayed

```{r message=FALSE, warning=FALSE}
toplotNational <- d[,.(n=sum(value)),keyby=.(age,Date)]
toplotNational[,level:="all"]
toplotNational[,geo:="all"]
toplotDistrict <- d[,.(n=sum(value)),keyby=.(age,district,Date)]
toplotDistrict[,level:="district"]
setnames(toplotDistrict,"district","geo")
toplotVillage <- d[,.(n=sum(value)),keyby=.(age,village,Date)]
toplotVillage[,level:="village"]
setnames(toplotVillage,"village","geo")
tosave <- rbind(toplotNational,toplotDistrict,toplotVillage)
tosave[,pop:=10000]

#readr::write_lines(jsonlite::toJSON(tosave),path="json/epicurve-by-day-dodged-by-age.json")

q <- ggplot(toplotNational,aes(x=Date,y=n,fill=age))
q <- q + geom_col(pos="dodge",width=0.5)
q <- q + labs(title=sprintf("Daily epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
#q <- q + scale_x_discrete("Epiweek")
q <- q + scale_y_continuous("Number of reported cases")
q <- q + scale_fill_brewer("Age",palette="Set1")
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```


## Age and sex distribution over different time frames

- We display the number of cases, split by age/sex on the x-axis
- We need the ability to display different time frames (e.g. per week, last week, over multiple weeks)

```{r message=FALSE, warning=FALSE}
toplotNational <- d[,.(n=sum(value)),keyby=.(age,sex,epiWeek)]
toplotNational[,level:="all"]
toplotNational[,geo:="all"]
toplotDistrict <- d[,.(n=sum(value)),keyby=.(age,sex,district,epiWeek)]
toplotDistrict[,level:="district"]
setnames(toplotDistrict,"district","geo")
toplotVillage <- d[,.(n=sum(value)),keyby=.(age,sex,village,epiWeek)]
toplotVillage[,level:="village"]
setnames(toplotVillage,"village","geo")
tosave <- rbind(toplotNational,toplotDistrict,toplotVillage)
tosave[,pop:=10000]

#readr::write_lines(jsonlite::toJSON(tosave),path="json/age-and-sex-dist-over-different-time.json")

toplot <- toplotNational[epiWeek==max(epiWeek)]
toplot[,var:=sprintf("%s, %s",age,sex)]

q <- ggplot(toplot,aes(x=var,y=n))
q <- q + geom_col(alpha=0.7)
q <- q + labs(title=sprintf("Age and sex distribution %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
q <- q + scale_x_discrete("")
q <- q + scale_y_continuous("Number of reported cases")
#q <- q + facet_grid(.~sex)
#q <- q + coord_flip()
q
```

## Weekly epicurves by age/sex

Here we display four weekly epicurves, one for each age/sex combination.

Important to note:

- We display `year-isoweek` on the x-axis
- Weeks with zero cases must be displayed
- Y-axis remains the same height for all panels, to allow for easy comparison

```{r message=FALSE, warning=FALSE}
toplot <- d[,.(n=sum(value)),by=.(epiWeek,sex,age)]

q <- ggplot(toplot,aes(x=epiWeek,y=n))
q <- q + geom_col()
q <- q + labs(title=sprintf("Weekly epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
q <- q + scale_x_discrete("Epiweek")
q <- q + scale_y_continuous("Number of reported cases")
q <- q + facet_grid(sex~age)
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```

## Weekly epicurves by geographical area

Here we display multiple weekly epicurves, one for each geographical area.

Important to note:

- We should probably be able to choose the granularity of geographical area (region/district/village)
- We display `year-isoweek` on the x-axis
- Weeks with zero cases must be displayed
- Y-axis remains the same height for all panels, to allow for easy comparison (this should probably be a toggle?)
- Very important: We should also implement one version where the outcome is: `(number of reported cases)/(estimation population)*10000` (i.e. number of reported cases per 10.000 population).

```{r message=FALSE, warning=FALSE}
toplot <- d[,.(n=sum(value)),by=.(epiWeek,district)]

q <- ggplot(toplot,aes(x=epiWeek,y=n))
q <- q + geom_col(alpha=0.7)
q <- q + labs(title=sprintf("Epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
q <- q + scale_x_discrete("Epiweek")
q <- q + scale_y_continuous("Number of reported cases")
q <- q + facet_wrap(~district)
q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```

## Map by geographical area

Here we display a map with categorized number of cases.

Important to note:

- We should probably be able to choose the granularity of geographical area (region/district/village)
- We should be able to change the time-frame
- Not reporting regions should be highlighted
- The graphing/outcome should be categorical NOT a continuous gradient. Probably no more than 4 categories.
- Very important: We should also implement one version where the outcome is: `(number of reported cases)/(estimation population)*10000` (i.e. number of reported cases per 10.000 population).

```{r message=FALSE, warning=FALSE}
toplot <- d[epiWeek==max(epiWeek),.(n=sum(value)),by=.(epiWeek,district)]
m <- merge(map,toplot,by.x="id",by.y="district",all.x=T)
b <- unique(quantile(m$n,na.rm=T))
m[,fillVal:=cut(n,breaks = b)]
l <- levels(m$fillVal)
m[,fillVal:=as.character(fillVal)]
m[is.na(fillVal),fillVal:="Not reporting"]
m[,fillVal:=factor(fillVal,levels=c("Not reporting",l))]

q <- ggplot()
q <- q + geom_map(data = m, aes(x = long, y = lat, map_id = id, fill=fillVal),
map = m, colour = "black", size=0.1)
q <- q + coord_map("albers",  at0 = 45.5, lat1 = 29.5)
q <- q + scale_fill_manual("",values=c("grey","#ffffb2","#fecc5c","#fd8d3c","#e31a1c"),drop=F)
q <- q + scale_x_continuous("")
q <- q + scale_y_continuous("")
q <- q + labs(title="Reported cases in epiweek X")
q
```

## Barcharts by district

This is very similar to the above map, but allows for a more nuanced view of the numbers.

Important to note:

- We should probably be able to choose the granularity of geographical area (region/district/village)
- We should be able to change the time-frame
- The graphing/outcome should be CONTINUOUS
- Very important: We should also implement one version where the outcome is: `(number of reported cases)/(estimation population)*10000` (i.e. number of reported cases per 10.000 population).

```{r message=FALSE, warning=FALSE}
toplot <- d[epiWeek==max(epiWeek),.(n=sum(value)),by=.(epiWeek,district)]

q <- ggplot(toplot,aes(x=district,y=n))
q <- q + geom_col(alpha=0.7)
q <- q + labs(title=sprintf("Epicurve from %s to %s",minDate,maxDate))
q <- q + labs(caption=sprintf("Created %s",lubridate::today()))
q <- q + scale_x_discrete("")
q <- q + scale_y_continuous("Number of reported alerts")
q <- q + coord_flip()
q <- q + labs(title="Cases in epiweek X")
#q <- q + theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=0.5))
q
```

## District/Person reporting funnel plot A

The idea of this funnel plot is to identify districts/people who are reporting worse than expected.

For each month, we count the number of messages sent, and the number of correctly sent messages. From this, we generate an "expected proportion of received messages that are correct" (e.g. 80%). Then, for i = 1, ..., 100 (or higher, as necessary) we calculate the 2.5th and 97.5th percentiles according to the binomial distribution. That is, (e.g. for i=40) what is the 2.5th and 97.5th percentile of a Binom(n=40, p=0.8) distribution. These percentiles are our boundaries as displayed in the graph.

Important to note:

- We should be able to switch between district/people/other grouping measure
- We should be able to change the time-frame
- Maybe only the people/groups who are "lower than expected" should be highlighted in some way?

```{r message=FALSE, warning=FALSE}

lastTwoEpiMonths <- rev(sort(unique(dWide$epiMonth)))[1:2]

reporters <- dWide[epiMonth %in% lastTwoEpiMonths,.(
  reporters=length(unique(datacollector)),
  reported=max(isMessage),
  messages=sum(isMessage),
  messagesValid=sum(isMessageValid)
  ),keyby=.(
  epiMonth,
  district
  )]
reporters[,epiMonthOrdered:=factor(epiMonth,levels=rev(sort(unique(epiMonth))))]

reporters[,propValid:=sum(messagesValid)/sum(messages),by=.(epiMonth)]
reporters[,threshold025:= qbinom(0.025, messages, propValid)]
reporters[,threshold975:= qbinom(0.975, messages, propValid)]

thresholdsBase <- unique(reporters[,c("epiMonth","epiMonthOrdered","propValid")])
thresholds <- vector("list",length=nrow(thresholdsBase))
for(i in 1:nrow(thresholdsBase)){
  thresholds[[i]] <- data.table(
    messages=1:max(reporters$messages+5),
    propValid=thresholdsBase[i]$propValid,
    epiMonth=thresholdsBase[i]$epiMonth,
    epiMonthOrdered=thresholdsBase[i]$epiMonthOrdered
    )
}
thresholds <- rbindlist(thresholds)
thresholds[,threshold025:= qbinom(0.025, messages, propValid)]
thresholds[,threshold975:= qbinom(0.975, messages, propValid)]

q <- ggplot(mapping=aes(x=messages))
q <- q + geom_ribbon(data=thresholds,mapping=aes(ymin = -Inf, ymax = threshold025/messages, fill = "0low"), alpha = 0.4)
q <- q + geom_ribbon(data=thresholds,mapping=aes(ymin = threshold025/messages, ymax = threshold975/messages, fill = "1normal"), alpha = 0.4)
q <- q + geom_ribbon(data=thresholds,mapping=aes(ymin = threshold975/messages, ymax = Inf, fill = "2high"), alpha = 0.4)
q <- q + geom_point(data=reporters,mapping=aes(y=messagesValid/messages))
q <- q + ggrepel::geom_label_repel(data=reporters,mapping=aes(y=messagesValid/messages,label=district),alpha=0.4)
q <- q + facet_wrap(~epiMonthOrdered, ncol=1)
q <- q + scale_fill_manual("",values=c("red","blue","yellow"),labels=c("Lower than expected","As expected","Higher than expected"))
q <- q + guides(fill = guide_legend(reverse=T))
q <- q + scale_x_continuous("Number of messages received")
q <- q + scale_y_continuous("Percent of received messages that were valid",labels=scales::percent)
q
```

## District/Person reporting funnel plot B

The idea of this funnel plot is to identify districts/people who are reporting worse than expected.

For each month, we count the number of messages sent, and the number of correctly sent messages. From this, we generate an "expected proportion of received messages that are correct" (e.g. 80%). Then, for each district/person, we calculate the 2.5th and 97.5th percentiles according to the binomial distribution. That is, (e.g. if Oslo sent 40 messages) what is the 2.5th and 97.5th percentile of a Binom(n=40, p=0.8) distribution. These percentiles are our boundaries as displayed in the graph.

Important to note:

- We should be able to switch between district/people/other grouping measure
- We should be able to change the time-frame
- Maybe only the people/groups who are "lower than expected" should be highlighted in some way?


```{r}
q <- ggplot(reporters[epiMonth %in% lastTwoEpiMonths],aes(x=reorder(district,messagesValid/messages),y=messagesValid/messages))
#q <- ggplot(x,aes(x=reorder(priority1_cat,-perc),y=perc))
#q <- q + geom_col(alpha=0.0)
q <- q + geom_linerange(mapping=aes(ymin=0,ymax=threshold025/messages),lwd=4,alpha=0.6,colour="red")
q <- q + geom_linerange(mapping=aes(ymin=threshold025/messages,ymax=threshold975/messages),lwd=4,alpha=0.6,colour="blue")
q <- q + geom_linerange(mapping=aes(ymin=threshold975/messages,ymax=1),lwd=4,alpha=0.6,colour="yellow")
q <- q + scale_y_continuous("Percent of received messages that were valid",labels = scales::percent_format(),lim=c(0,1))
q <- q + geom_point(size=5)
q <- q + scale_x_discrete("")
q <- q + coord_flip()
q <- q + facet_wrap(~epiMonthOrdered, ncol=1)
q
```













